---
title: "Developing Data Products- Project's Part 2 -Reproducible Pitch Presentation"
author: "tonuta"
date: "May 14, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Slide 2

Developing an App: Relationship between variables and MPG based on mtcars dataset.

Shiny Application:

URL: https://tonuta.shinyapps.io/Developing-Data-Products---Part-1/

A shiny-App shows the relationship between variables and miles per gallon (MPG).

Get all SourceCode @Github:
https://github.com/tonuta/Developing-Data-Products---Part-1

## Slide 3

mtcars dataset - Description (from Motor Trend Car Road Tests)

Source of data (Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391-411.)

The data comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

```{r, echo = TRUE, warning = FALSE, message = FALSE}
require(plotly)
library(datasets)
head(mtcars, 3)
```

## Slide 4
mtcars dataset - A data frame with 32 observations on 11 variables.

- [, 1]	mpg	Miles/(US) gallon
- [, 2]	cyl	Number of cylinders
- [, 3]	disp	Displacement (cu.in.)
- [, 4]	hp	Gross horsepower
- [, 5]	drat	Rear axle ratio
- [, 6]	wt	Weight (lb/1000)
- [, 7]	qsec	1/4 mile time
- [, 8]	vs	V/S
- [, 9]	am	Transmission (0 = automatic, 1 = manual)
- [,10]	gear	Number of forward gears. - [,11]	carb	Number of carburetors.

## Slide 5

```{r, echo = TRUE, warnings = FALSE, message = FALSE}
plot_ly(mtcars, x = ~wt, y = ~mpg, type = "scatter",
        color = ~factor(cyl), size = ~hp)
```       

## Slide 6

Analysis - main code

  formulaTextPoint <- reactive({
    paste("mpg ~", "as.integer(", input$variable, ")")  })

  fit <- reactive({
    lm(as.formula(formulaTextPoint()), data=mpgData)  })
  ...
  output$fit <- renderPrint({
    summary(fit()) })

  output$mpgPlot <- renderPlot({
    with(mpgData, {
      plot(as.formula(formulaTextPoint()))
      abline(fit(), col=2)
    })  })
